Classical probability measures as pullback of quantum probability measures?

### 2022-05-03 pullback.tex
pullback.lyx The version attached "On Wed, May 4, 2022 at 4:43 AM Yu-Tsung Tai <yuttai@alumni.iu.edu>"

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will happen. To maintain the coherence,
$\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\Omega\backslash E\right)=\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\Omega\backslash E$ is the complement of $E$, and we may
simply write $\overline{E}$ if the sample space~$\Omega$ is clear
from the context.  Moreover, we require $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

Since the previous abstraction doesn't specify the process to generate
the outcomes, this process could well be a quantum experiment. Let
us prepare a beam of one kind of spin~$1$ particles, and split it
by an inhomogeneous magnetic field along the $z$-axis as a well-known
Stern-Gerlach type experiment \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum}.
This beam will be split into three beams, and we will call the outcome
beams bent up as $\ket{0}$, bent down as $\ket{1}$, and not-bent
as $\ket{2}$. Instead of sending a beam of particles, if only one
particle is sent to the beam splitter, we will only get one of the
outcomes $\ket{0}$, $\ket{1}$, and $\ket{2}$, and there is a probability
to get each outcome. In the language of our abstraction, the set of
outcomes is $\Omega_{0}=\{\ket{0},\ket{1},\ket{2}\}$. All possible
events are $\emptyset$, $\left\{ \ket{0}\right\} $, $\left\{ \ket{1}\right\} $,
$\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{0}$. Each time we perform this experiment, it induces
a classical probability measure $\mu_{0}\colon2^{\Omega_{0}}\rightarrow\left[0,1\right]$.
Moreover, if we prepare the particles in the ``same'' way as the
particle inducing $\mu_{0}$, they will induce the same classical
probability measure $\mu_{0}$.

>>>>

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will happen. To maintain the coherence,
$\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\Omega\backslash E\right)=\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\Omega\backslash E$ is the complement of $E$, and we may
simply write $\Omega\backslash E$ as $\overline{E}$ if the sample
space~$\Omega$ is clear from the context.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

Since the previous abstraction doesn't specify the process to generate
the outcomes, this process could well be a quantum experiment. Let
us prepare a beam of one kind of spin~$1$ particles, and split it
by an inhomogeneous magnetic field along the $z$-axis as a well-known
Stern-Gerlach type experiment \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum,schwinger2001quantum}.
This beam will be split into three beams, and we will call the outcome
beams bent up as $\ket{0}$, bent down as $\ket{1}$, and not-bent
as $\ket{2}$. Instead of sending a beam of particles, if only one
particle is sent to the beam splitter, we will only get one of the
outcomes $\ket{0}$, $\ket{1}$, and $\ket{2}$, and there is a probability
to get each outcome. In the language of our abstraction, the set of
outcomes is $\Omega_{0}=\{\ket{0},\ket{1},\ket{2}\}$. All possible
events are $\emptyset$, $\left\{ \ket{0}\right\} $, $\left\{ \ket{1}\right\} $,
$\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{0}$. Each time we perform this experiment, it induces
a classical probability measure $\mu_{0}\colon2^{\Omega_{0}}\rightarrow\left[0,1\right]$.
Moreover, if we prepare the particles in the ``same'' way as the
particle inducing $\mu_{0}$, they will induce the same classical
probability measure $\mu_{0}$.
